Learning by Teaching a Robot: the Case of Handwriting

### 2015-07-07 main.tex
More work on technical implementation

The development of the necessary components for embedding the handwriting
learning algorithm presented in Section \ref{sec:learningAlgorithm} in the
humanoid agent are presented in the sections which follow.

\begin{figure}[ht]
\centering

@@

\caption{\small Overview of the system. In total, the system runs about 10 ROS nodes,
    distributed over the robot itself, a central laptop and Android tablets. Since
    the system does not require any CPU-intensive process, the laptop could be
    removed and the whole logic could run on the robot. Due to the relative
    difficulty to deploy and debug ROS nodes directly on the robot, we however
    relied on the laptop during our experiments.}

@@

@@

\subsubsection{Robot Trajectory Following Movements}

Using simulated handwriting provides an opportunity for the robot's writing to
appear smoother than would be achievable with a writing instrument. However, the
robot's motions must still sufficiently match the displayed trajectory in order
capture the engagement of the child participant in the action. Aldebaran's NaoQi
API is used for the inverse kinematics of the trajectory following. The Robot
Operating System (ROS)\footnote{The ROS stack for {\sc nao} is available at:
\url{http://wiki.ros.org/nao_robot}.} is used for integration of the {\sc nao}
with external reference frames, such as the tablet's location, using the $tf$
transformation library \cite{Foote2013}.

When using simulated handwriting, it is no longer necessary that the robot
engages in the typical style of handwriting of using a writing instrument at a
desk.  Having the robot point at a vertical writing surface to cause the
trajectory to appear (as in Figure \ref{fig:naoWriting}) has several advantages:

@@

    \item The working space of the robot increases, both in the technical sense
        and the interaction sense: someone can, in theory, show the tablet to
        the robot from across the room and have it still respond, without
        needing the tablet to be within arm's reach.

    \item Concerns about whether or not the child would start mimicking the
        robot's incorrect writing form (\eg pen grip) are mitigated.

@@

\end{itemize}

\begin{figure}[thpb]
     \begin{center}
            \includegraphics[width=0.7\linewidth]{naoWriting2.png}
    \end{center}
    \caption{\small A demonstration of the robot simulating the writing of a word with
    its finger. The motion of the robot is synchronised with the display of the
    tablet, communicating over ROS.}%

   \label{fig:naoWriting}
\end{figure}

>>>>

\begin{figure}[ht!]
\centering

@@

\caption{\small Overview of the system. In total, the system runs about 10 ROS nodes,
    distributed over the robot itself, a central laptop and Android tablets.}

@@

The overall architecture of the system (Figure~\ref{fig:archi}) is therefore
spread over several devices: the {\sc nao} robot itself, that we address via
both a ROS API\footnote{The ROS stack for {\sc nao} is available at
\url{http://wiki.ros.org/nao_robot}.} and the Aldebaran-provided NaoQI API, one
to four Android tablets (the main tablet is used to print the robot's letter and
to acquire the children' demonstrations. More tablets have been used in some
studies, either to let the child input words to be written, or for the
experimenter to qualitatively annotate the interaction in a synchronized
fashion), and a central laptop running the machine learning algorithms, the
robot's handwriting gesture generation and high level control of the activity.

Since the system does not actually require any CPU-intensive process, the laptop
could be removed and the whole logic could run on the robot. Due to the relative
difficulty to deploy and debug ROS nodes directly on the robot, we however
effectively kept the laptop during our experiments.

Most of the nodes are written in Python, and the whole source code of the
project is available online\footnote{The primary repository is
\url{https://github.com/chili-epfl/cowriter_letter_learning}.}

@@

\subsubsection*{Trajectory Following Movements}

Using simulated handwriting provides an opportunity for the robot's writing to
appear smoother than would be achievable with a writing instrument. However, the
robot's motions must still sufficiently match the displayed trajectory in order
capture the engagement of the child participant in the action. Aldebaran's NaoQi
API is used for the inverse kinematics of the trajectory following. ROS is used
for integration of {\sc nao} with external reference frames, such as the
tablet's location, using the $tf$ transformation library~\cite{Foote2013}.

When using simulated handwriting, it is no longer necessary that the robot
engages in the typical style of handwriting of using a writing instrument at a
desk. Having instead the robot point at the writing surface to cause the
trajectory to appear (as seen in Figure \ref{fig:schools}, left) has several advantages:

@@

    \item The working space of the robot increases, both in the technical sense
        and the interaction sense: someone can, in theory, show the tablet to
        the robot from across the room and have it still respond, without
        needing the tablet to be within arm's reach,

    \item Concerns about whether or not the child would start mimicking the
        robot's incorrect writing form (\eg pen grip) are mitigated,

@@

\end{itemize}
%
%\begin{figure}[thpb]
%     \begin{center}
%            \includegraphics[width=0.7\linewidth]{naoWriting2.png}
%    \end{center}
%    \caption{\small A demonstration of the robot simulating the writing of a word with
%    its finger. The motion of the robot is synchronised with the display of the
%    tablet, communicating over ROS.}%
%
%   \label{fig:naoWriting}
